Empirical Analysis of the Role of Energy in Economic Growth

### 2013-01-19 EconGrowth.Rnw
Now includes a function to do Cobb-Douglas with energy.

<<testing>>=
mat <- round(matrix(c(0.9, 0.89, 200, 0.045, 2.0), c(1, 5)), 4)
rownames(mat) <- "$y_{t-1}$"
# colnames(mat) <- c("$R^2$", "$\\bar{R}^2$", "F-stat", "S.E.E", "DW")
colnames(mat) <- c("$\\lambda$", "$\\alpha$", "$\\beta$", "$\\gamma$", "DW")
mat <- xtable(mat)
@

<<Test Print Sanitized Table, results='asis'>>=
print(mat, sanitize.text.function = function(x){x})
@

@@

<<cobb-douglass>>=
# Establish guess values for alpha and lambda.
lambdaGuess <- 0.0 # guessing lambda = 0 means there is no technological progress.
alphaGuess <- 0.3 # a typical value for alpha, the coefficient on capital stock

@@

\section{Cobb-Douglas With Q}

<<cdq>>=
# Establish guess values for alpha, beta, and lambda.
# lambdaGuess <- 0.0 # guessing lambda = 0 means there is no technological progress.
# alphaGuess <- 0.2 # a typical value for alpha
# betaGuess <- 0.6 # a typical value for beta

# Runs a non-linear least squares fit to the data.
# modelCDQ <- nls(iGDP ~ exp(lambda*iYear) * iCapStk^alpha * iLabor^beta * iQ^(1.0 - alpha - beta), 
#                 start=(list(lambda=lambdaGuess,alpha=alphaGuess,beta=betaGuess)),
#                 data=dataTable)

# Reparameterize to ensure that we meet the constraint that alpha + beta + gamma = 1.0.
# 0 < a < 1
# 0 < b < 1
# alpha = min(a, b)
# beta = b - a
# gamma = 1 - max(a, b)

@@

<<CDq Results, results='asis'>>=
print(tableCDq, caption.placement="top")
# According to http://cran.r-project.org/web/packages/xtable/vignettes/xtableGallery.pdf, Section 3.1, I should 
# be able to use the "sanitize.text.function" parameter to allow markup in column headers. But this next
# line is not working at the present time. --MKH, 18 Jan 2012.
# print(tableCDq, sanitize.text.function = function(x){x})

>>>>

@@

<<cobb-douglas>>=
# Establish guess values for alpha and lambda.
lambdaGuess <- 0.0 # guessing lambda = 0 means there is no technological progress.
alphaGuess <- 0.3 # a typical value for alpha, the coefficient on capital stock

@@

\section{Cobb-Douglas With Energy}

<<cobbDouglasEnergy function>>=
###############################
# This function fits the Cobb Douglas production function including energy to historical data
# The functional form is
#
# iGDP = exp(lambda * itime) * iCapStk^alpha * iLabor^beta * iEnergy^gamma
# 
# iYear: time indexed to 0.0 at the first year [years since beginning of data set]
# iGDP = time series GDP data indexed to 1.0 at the first year [-]
# iCapStk = time series capital stock data indexed to 1.0 at the first year [-]
# iLabor = time series labor data indexed to 1.0 at the first year [-]
# iEnergy = time series energy data indexed to 1.0 at the first year [-]. 
#      This may be any of heat (iQ), exergy (iX), or useful work (iU).
# 
# returns a data object containing three rows and 5 columns.
#   col 1: lambda
#   col 2: alpha
#   col 3: beta
#   col 4: gamma
#   row 1: -95% CI for each parameter
#   row 2: estimate for each parameter
#   row 3: +95% CI for each parameter
#
# dataTable is the table to be used for this curve fit
# iEnergy is the name of the column to be used for energy.
#   This may be any of heat (iQ), exergy (iX), or useful work (iU).
##
cobbDouglasEnergy <- function(dataTable, iEnergy){
  # Reparameterize to ensure that we meet the constraints:
  # * alpha + beta + gamma = 1.0.
  # * alpha, beta, and gamma are all between 0.0 and 1.0.
  # To do this, we reparameterize as
  # * 0 < a < 1
  # * 0 < b < 1
  # * alpha = min(a, b)
  # * beta = b - a
  # * gamma = 1 - max(a, b)
  lambdaGuess <- 0.0 # guessing lambda = 0 means there is no technological progress.
  alphaGuess <- 0.3 # a typical value for alpha
  betaGuess <- 0.7 # a typical value for beta
  modelCDe <- nls(iGDP ~ exp(lambda*iYear) * 
                    iCapStk^min(a,b) * iLabor^abs(b-a) * 
                    iQ^(1.0 - max(a,b)), 
                  algorithm = "port",
                  start = list(lambda=lambdaGuess, a=alphaGuess, b=alphaGuess+betaGuess),
                  lower = list(lambda=-Inf, a=0, b=0),
                  upper = list(lambda=Inf, a=1, b=1),
                  data = dataTable)
  aicCDe <- AIC(modelCDe, k=2); aicCDe # Checks validity of the model. AIC stands for Akaike's Information Criterion

  summaryCDe <- summary(modelCDe); summaryCDe # Gives the nls summary table

  # Provides confidence intervals on lambda, a, and b. But, we need CIs on alpha and beta.
  ciCDe <- confint(modelCDe, level = 0.95); ciCDe

  a <- as.numeric(coef(modelCDe)["a"])
  b <- as.numeric(coef(modelCDe)["b"])
  lambda <- as.numeric(coef(modelCDe)["lambda"])
  alpha <- a
  beta <- b - a
  gamma <- 1.0 - alpha - beta

  # Report results with SE
  beta.est <- deltaMethod(modelCDe, "b-a"); beta.est # Reports results for beta, because beta = b - a.
  gamma.est <- deltaMethod(modelCDe, "1-b"); gamma.est # Reports results for gamma, because gamma = 1 - b.

  # Now calculate confidence intervals.
  dofCDe <- summaryCDe$df[2]; dofCDe # Gives the degrees of freedom for the model.
  tvalCDe <- qt(0.975, df = dofCDe); tvalCDe
  betaCICDe <- with(beta.est, Estimate + c(-1.0, 1.0) * tvalCDe * SE); betaCICDe # Gives the confidence interval on beta.
  gammaCICDe <- with(gamma.est, Estimate + c(-1.0, 1.0) * tvalCDe * SE); gammaCICDe # Gives the confidence interval on alpha.

  # Combine all estimates and their confidence intervals into data frames with intelligent row names
  estCDe <- data.frame(lambda = lambda, alpha = alpha, beta = beta, gamma = gamma); estCDe
  #row.names(estCDe) <- "Cobb-Douglas with e: $y = e^{\\lambda t}k^{\\alpha}l^{\\beta}e^{\\gamma}$"
  row.names(estCDe) <- "CobbDouglas with e"
  # The [1] subscripts pick off the lower confidence interval
  lowerCDe <- data.frame(lambda = ciCDe["lambda","2.5%"], alpha = ciCDe["a", "2.5%"], beta = betaCICDe[1], gamma = gammaCICDe[1])
  row.names(lowerCDe) <- "- 95% CI"
  # The [2] subscripts pick off the lower confidence interval
  upperCDe <- data.frame(lambda = ciCDe["lambda","97.5%"], alpha = ciCDe["a", "97.5%"], beta = betaCICDe[2], gamma = gammaCICDe[2])
  row.names(upperCDe) <- "+ 95% CI"

  # Now create the data for a table.
  dataCDe <- rbind(lowerCDe, estCDe, upperCDe); dataCDe

  return(dataCDe)
}
@

\subsection{Cobb-Douglas with Q}

<<cdq>>=

dataCDe <- cobbDouglasEnergy(dataTable, dataTable["iQ"])

tableCDe <- xtable(dataCDe, caption="U.S. 1980-2011.", digit = c(4, 4, 2, 2, 3))

print("********** Done with function call **************")

# Establish guess values for alpha, beta, and lambda.
# lambdaGuess <- 0.0 # guessing lambda = 0 means there is no technological progress.
# alphaGuess <- 0.2 # a typical value for alpha
# betaGuess <- 0.6 # a typical value for beta
# 
# Runs a non-linear least squares fit to the data.
# modelCDQ <- nls(iGDP ~ exp(lambda*iYear) * iCapStk^alpha * iLabor^beta * iQ^(1.0 - alpha - beta), 
#                 start=(list(lambda=lambdaGuess,alpha=alphaGuess,beta=betaGuess)),
#                 data=dataTable)
# 
# Reparameterize to ensure that we meet the constraint that alpha + beta + gamma = 1.0.
# 0 < a < 1
# 0 < b < 1
# alpha = min(a, b)
# beta = b - a
# gamma = 1 - max(a, b)

@@

<<CDq Results, results='asis'>>=

print(tableCDe, caption.placement="top")
print(tableCDq, caption.placement="top")
# According to http://cran.r-project.org/web/packages/xtable/vignettes/xtableGallery.pdf, Section 3.1, I should 
# be able to use the "sanitize.text.function" parameter to allow markup in column headers. But this next
# line is not working at the present time. --MKH, 18 Jan 2012.
# print(tableCDq, sanitize.text.function = function(x){x})
